### 2023-05-04 paper/model-documentation.qmd
updating model, model documentation

---
title: Modeling costs and benefits of genomic wastewater surveillance
subtitle: Model appendix draft
#author:
#  - name: Pedro Nascimento de Lima
#    email: plima@rand.org
#    affiliations: 
#        - id: rand
#          name: RAND Corporation
#          department: Engineering and Applied Sciences
#          address: 1776 Main Street
#          city: Santa Monica
#          state: CA
#          postal-code: 90401
#    attributes:
#        corresponding: true
#  - name: and colleagues
#    email: others@rand.org
#    affiliations:
#        - ref: rand
abstract: |
  Introduces a model to assess the value provided by wastewater genomic surveillance network. This document is meant to be used as a supplementary appendix of the paper.
keywords: 
  - Genomic Surveillance
  - Pandemic Response
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      #name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
  elsevier-html:
    keep-tex: true
    journal:
      #name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
  docx:
      cite-style: super
---

<!-- Template source: https://github.com/quarto-journals/elsevier download repository and place it within this folder.-->

# Model

We start with a standard disease transmission model describing person-to-person transmission of an infectious pathogen across $n$ jurisdictions. The state-space of the model is comprised of a set of health statuses $\mathbf{S}, \mathbf{E}, \mathbf{P}, \mathbf{I}, \mathbf{R}$, each a $1 \times n$ vector representing the population within each compartment, respectively representing the number of Susceptibles, Exposed, Pre-Symptomatic, Infected and Removed[^1]. @eq-ode1 describes a system of differential equations representing the model, and @tbl-parameters contains parameter definitions, values and their sources.

[^1]: We start with a simplified version including only a minimal set of disease states before expanding the state-space of the model to include asymptomatic, mildy symptomatic and severely symptomatic compartments

$$ 
\begin{align}
\dot{\mathbf{S}} &=\ - \boldsymbol{\lambda}_t \odot \mathbf{S} + p_S \mathbf{m}_S \\
\dot{\mathbf{E}} &=\ \boldsymbol{\lambda}_t \odot \mathbf{S} - \sigma \mathbf{E} + p_E \mathbf{m}_E \\
\dot{\mathbf{P}} &=\ \sigma \mathbf{E} - \delta \mathbf{P} + p_P \mathbf{m}_P \\
\dot{\mathbf{I}} &=\ \delta \mathbf{P} - \gamma \mathbf{I} + p_I \mathbf{m}_I \\
\dot{\mathbf{R}} &=\ \gamma \mathbf{I} + p_R \mathbf{m}_R \\
\end{align}
$$ {#eq-ode1}

where $\dot{\mathbf{S}}, \dot{\mathbf{E}}, \dot{\mathbf{P}}, \dot{\mathbf{I}}, \dot{\mathbf{R}}$ represent the time derivatives of each disease status, greek letters represent progression rates and $p_x\mathbf{m}_x$ terms represent migration flows $\mathbf{m}_x$ modulated by propensity to travel $p_x$ at disease state $x \in \{S, E, P, I, R\}$. Force of infection $\boldsymbol{\lambda}_t$ is computed using a time-varying effective contact-rate vector $\boldsymbol{\beta}_t$ which will depend on contact rates within each strata as well as nonpharmaceutical interventions (discussed later). Infectiousness of each disease status $\mathbf{P}$ and $\mathbf{I}$ are modulated by constants $c_p$ and $c_i$.

$$ 
\boldsymbol{\lambda}_t = \boldsymbol{\beta}_t (c_p\mathbf{P}_t + c_i\mathbf{I}_t)
$$ {#eq-force-inf}

## Nonpharmaceutical interventions {#npi}

Following mitigation plans adopted during the COVID-19 pandemic, nonpharmaceutical interventions (NPIs) are modeled as a set of intervention portfolios organized within intensity levels with a pre-specified marginal effectiveness $\tau$ at each intervention level $l \in L = \{1,2,3,...,l_{max}\}$. In this analysis, $\tau$ is specified as a percent reduction in infectious contacts induced by a set of interventions introduced by public health decision-makers. In principle, $\tau_l$ can be specified so that it is dependent on the intervention level as in [@nascimentodelima2021], but this analysis assumes a constant, additive marginal effectiveness of interventions. For instance, if $\tau = 0.15$, then introducing the first intervention level reduces potential infectious contacts by 15% relative to a no-intervention scenario and introducing the second intervention level reduces infectious contacts by 30%. Under this specification, $l_{max} \tau \in (0,1)$.

$$ 
\boldsymbol{\beta}_t = (1-\mathbf{l}_t\tau) \boldsymbol{\beta}_b
$$ {#eq-betat}

and the intervention level $\mathbf{l}_t$ is a jurisdiction-dependent nonpharmaceutical intervention level. $\mathbf{l}_t$ can be either specified as an exogenous variable or an endogenous variable controlled by local publich health departments.

## Introducing and removing NPIs

In this model, local public health decision-makers follow epidemiological outcomes that are proportional to $\mathbf{I_t}$ (such as case counts) to decide the appropriate level of intervention. Thus, we specify a target NPI level $\mathbf{l}_t^*$ as a function of the disease state.

$$ 
\mathbf{l}_t^* = cI
$$ {#eq-targetnpi}

This formulation approximates phased mitigation plans where policymakers introduce or remove interventions depending on disease prevalence. While public health decision-makers strive to use up-to-date information to introduce and remove NPIs, they do not have perfect and instantaneous information about the disease state. In continuous time, we represent this process with a stock vector $\mathbf{L}$ representing the NPI level in effect

$$ 
\dot{\mathbf{L}} = \frac{1}{a}(\mathbf{l}_t^*-\mathbf{L})
$$ {#eq-actualnpi}

where $a$ represents the adjustment time for NPIs (i.e., one or two weeks) using conventional measures.

## Genomic surveillance

Genomic surveillance might provide an early signal to decision makers by providing information about current disease prevalence, hence effectively reducing $a$. One might also represent genomic surveillance using a mechanistic formulation where

```{=html}
<!-- 
Sections to complete

## Travel

Individuals travel from jurisdiction $i$ to $j$ at a daily rate $r_{ij}$. This baseline rate is multiplied by a factor  by a factor  The daily travel flows $t_{ij} \sim Pois(\lambda_{ij})$. We used data from {blank} to estimate travel flows between

### Information sharing and coordination

Jurisdictions can share information and use each other's epidemiological indicators to decide when to introduce and remove Nonpharmaceutical interventions.

## Effect of genomic surveillance

Genomic surveillance is assumed to provide benefits to public health decision makers by providing an early signal of surges that would be otherwise detected later via testing of symptomatic individuals.

# Model parameters {#parameters}

@tbl-parameters presents the set of model parameters used in this analysis, their meaning, and their baseline values and sources used.

-->
```
# Tables

```{r}
#| label: tbl-parameters
#| tbl-cap: Parameters and values
#| echo: false

params <- data.frame(Parameter = "$\\lambda_{ij}$ Travel rate", `Value` = "Mobility Matrix A", Source = "[@Feynman1963118]")

knitr::kable(params)
```

# References {.unnumbered}