### 2024-02-19 reproduce-results.qmd
supporting variant strains

---
title: "results"
format: html
author: "Pedro Nascimento de Lima"
---

# Reproducibility notes

This quarto document reproduces the results section in Nascimento de Lima et al. (2024). *The value of environmental surveillance for pandemic response*. For details on how to install dependencies, and reproduce these results, see the `README.md` file.

# Run Results

First, we instantiate the model and run it over the experimental designs used in the paper.

```{r}

```

# Results

## Value of EWS under base-case results
{Describe dynamics and results uder base-case assumptions}

D. EWS provide meaningful value under base-case assumptions

## Conditions for high-value early warning 
{Describe conditions under which value is lower or higher}
D. Value of EWS hinges on timeliness of and adherence to NPIs. 
There was a

## EWS and policy coordination
D. EWS alleviate lack of policy coordination. 
There is a concern that EWS data will need to be integrated to be useful. However, we find that policy coordination and EWS act as substitutes. Policy coordination h

## Marginal Value of EWS
D. EWS deployment also face decreasing marginal returns
The value of EWS decreases as coverage increases.

# Results

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).